---
title: |
  <center>
  <strong>AKSTA Statistical Computing – Case Study 1</strong><br>
  
  Arthur Wunder (12431488), Temur ... (...), Valentina ... (...)<br>
  
  25 March 2025
  </center>
output: html_document
---

# Exercise 1

### Task a

```{r}
# Function with for-Loop
fib_ratio_for <- function(n) {
  fib <- numeric(n + 1)
  fib[1] <- 1
  fib[2] <- 1
  for (i in 3:(n + 1)) {
    fib[i] <- fib[i - 1] + fib[i - 2]
  }
  return(fib[2:(n + 1)] / fib[1:n])
}

# Function with while-Lopp
fib_ratio_while <- function(n) {
  fib <- numeric(n + 1)
  fib[1] <- 1
  fib[2] <- 1
  i <- 3
  while (i <= n + 1) {
    fib[i] <- fib[i - 1] + fib[i - 2]
    i <- i + 1
  }
  return(fib[2:(n + 1)] / fib[1:n])
}

```

### Task b

```{r}
# Benchmark with microbenchmark
library(microbenchmark)

n1 <- 200
n2 <- 2000

bench_results <- microbenchmark(
  for_200 = fib_ratio_for(n1),
  while_200 = fib_ratio_while(n1),
  for_2000 = fib_ratio_for(n2),
  while_2000 = fib_ratio_while(n2),
  times = 10L
)

print(bench_results)

```

The implementation using the for loop is faster than the one using while for both n = 200 and n = 2000, based on the benchmark results.

### Task c

```{r}
# Plot for n = 100
n <- 100
ratios <- fib_ratio_for(n)

plot(1:n, ratios, type = "l", col = "blue", lwd = 2,
     main = "Fibonacci Ratio Sequence r[i] = F[i+1] / F[i]",
     xlab = "i", ylab = "Ratio F[i+1] / F[i]")
abline(h = (1 + sqrt(5)) / 2, col = "red", lty = 2)  # Golden ratio
legend("topright", legend = c("Fibonacci ratio", "Golden ratio"), 
       col = c("blue", "red"), lty = c(1, 2), bty = "n")

```


The Fibonacci ratio sequence starts to stabilize around n ≈ 10–15.
From that point onward, the ratio fluctuates only very slightly.

The sequence converges to the golden ratio, which is approximately 1.6180
This is clearly visible in the plot where the blue line approaches and stays very close to the red dashed line representing the golden ratio.

# Exercise 2

### Task a

```{r}
# Function to compute rho_n using the gamma function
rho_gamma <- function(n) {
  num <- gamma((n - 1) / 2)
  denom <- gamma(1 / 2) * gamma((n - 2) / 2)
  rho <- num / denom
  return(rho)
}

```

### Task b

```{r}
# Evaluate rho_n for n = 2000
rho_n_2000 <- rho_gamma(2000)
print(rho_n_2000)


```

When computing ρₙ for large values like n = 2000 using the gamma() function, the result is NaN. This happens because the gamma function grows very rapidly, and evaluating gamma((n - 1)/2) or gamma((n - 2)/2) leads to numerical overflow. Since R cannot represent such large values accurately, the division results in undefined or infinite values, causing the final result to be NaN.

### Task c

```{r}
# Numerically stable version using logarithms
rho_gamma_stable <- function(n) {
  log_num <- lgamma(n / 2)
  log_denom <- lgamma(1 / 2) + lgamma((n - 1) / 2)
  log_rho <- log_num - log_denom
  return(exp(log_rho))
}


# Test for large n
rho_stable_2000 <- rho_gamma_stable(2000)
print(rho_stable_2000)


```

### Task d

```{r}

n_vals <- 10:2000
rho_vals <- sapply(n_vals, rho_gamma_stable)
y_vals <- rho_vals / sqrt(n_vals)

plot(n_vals, y_vals, type = "l", col = "darkgreen", lwd = 2,
     main = expression(paste("Plot of ", rho[n], " / ", sqrt(n))),
     xlab = "n", ylab = expression(rho[n] / sqrt(n)))


```

Not sure if this is correct... 

